---
title: "Introduction to FORESEE"
author: "Lisa-Katrin Turnhoff and Ali Hadizadeh Esfahani"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## FORESEE: a Tool for the Systematic Comparison of Translational Drug Response Modeling Pipelines


uniFied translatiOnal dRug rESponsE prEdiction platform FORESEE is a versatile open source software, implemented as R-package, that is designed to act as a scaffold in developing and benchmarking such computational drug response models. FORESEE not only introduces a uniform data format for public cell line and patient data sets, but also establishes a standardized environment for drug response prediction pipelines, incorporating state of the art preprocessing methods, model training algorithms and different validation techniques. A modular implementation of the different elements of the modeling pipeline facilitates a straightforward development of different combinatorial models, which can be used to re-evaluate and improve already existing modeling pipelines, as well as to develop and benchmark new ones.

We recommend to install <a href="https://github.com/JRC-COMBINE/FORESEEData">FORESEEData</a> package alongside this package.

### Table of Contents
**[Installation Instructions](#installation-instructions)**<br>
**[Usage Instructions](#usage-instructions)**<br>
**[Troubleshooting](#troubleshooting)**<br>
**[Compatibility](#compatibility)**<br>
**[Notes and Miscellaneous](#notes-and-miscellaneous)**<br>
**[Next Steps, Credits, Feedback, License](#next-steps)**<br>

## Installation Instructions
### Installing via Devtools (Recommended method):
Easiest way to install FORESEE is via <a href="https://cran.r-project.org/web/packages/devtools/">Devtools</a>.
After installing Devtools from cran, you can install FORESEE by:
```r
devtools::install_github(repo = "JRC-COMBINE/FORESEE")
```

### Alternative installation methods (Manual download):
...
```Shell
cd [Your desired directory]
git clone https://github.com/JRC-COMBINE/FORESEE.git
```

```Shell
R -e "devtools::install_local('./FORESEE/')"
```

## Usage Instructions
## ...
## ...
### ...
## ...
### Integrating User-defned Functions into FORESEE Pipeline
FORESEE is designed to be easily expandable. While there are numerous different pipelines 
already implemented in the package, user can conveniently change each part of the pipeline 
with their own method. To be exact, in ForeseeTrain CellResponseTransformation, DuplicationHandling,
HomogenizationMethod, GeneFilter, FeaturePreprocessing and BlackBox can be defined by the user. In this
section we demonstrate this capability of FORESEE through some examples.

#### User-defined CellResponseTransformation:
The underlying distribution of 'Response' has a great impact on the fitted model using a blackbox. Here,
the 'Response' is the cell line response to a drug, quantified by different measures, e.g. IC50. The 
measured IC50s are not necessarily distributed for an optimum model fitiing, so it is quite common in
preprocessing stage to 'transform' these Response values, for example by power transforming or changing 
to logarithmic scale, which are already implemented in FORESEE and can be easily used by CellResponseTransformation="powertransform" and CellResponseTransformation="logarithm" respectively.

But It is possible to implement your own transformation and pass your function 
CellResponseTransformation=User_Transformation to be used in the FORESEE pipeline.

FORESEE expects your transformation to be function which has one vector of numerics as input, and returns
one vector of numerics.

As an example, we implement a transformation that uniforms the Response and use it in FORESEE:
```r
Uniformer <- function(x){
    sort(runif(n = length(x), min = min(x), max = max(x)))[rank(x)]
}
```

Now we just have to pass Uniformer function to ForeseeTrain as the CellResponseTransformation:
```r
EnvForRunningUniformer <- new.env()
EnvForRunningUniformer$Uniformer <- Uniformer()
tmp <- with(EnvForRunningUniformer, {ForeseeTrain(TrainObject = GDSC,
                                                TestObject = GSE6434, DrugName = "Docetaxel",
                                                DuplicationHandling = "first", GeneFilter = "variance",
                                                CellResponseTransformation = Uniformer);
  ForeseeTest(TestObject = TestObject, ForeseeModel = ForeseeModel)}
)
```


# Troubleshooting
# Compatibility
# Notes and Miscellaneous
# Building the Extension Bundles
# Next Steps, Credits, Feedback, License


------------------------------------------------------------------------

## Default Parts Made by DEVTOOLS, for Reference, Will Be Removed Later
Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
